---
title: "lab1"
author: "Olivia Isbell"
date: "1/7/2022"
output: html_document
---

```{r setup, include=TRUE, messages = FALSE , warnings = FALSE }
knitr::opts_chunk$set(echo = TRUE, messages = FALSE , warnings = FALSE)
#attach packages 

library(tidyverse)
library(here)
library(sf)
library(tmap)
```

### Reading in the data 
```{r}
sf_trees <- read_csv(here("data" , "sf_trees" , "sf_trees.csv"), 
                     show_col_types = FALSE)

```


### Part 1: Wangling data and ggplot review 

**Example 1:** Find counts of observation by `legal_status` and wrangle a bit  
```{r}
### method 1 
sf_trees %>% 
  group_by(legal_status) %>% 
  summarize(tree_cout = n()) #counts number of rows 

### method 2 
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>%  #take of na in column
  rename(tree_count = n) %>%  
  relocate(tree_count) %>% #brings to the front 
  slice_max(tree_count, n = 5) %>% 
  arrange(desc(tree_count)) #can also use - for desc or none for lowest to highest
```

Make a graph 
```{r}
ggplot(data = top_5_status, aes(x= fct_reorder(legal_status, tree_count), y = tree_count)) +
  geom_col(fill = "darkgreen")+ 
  coord_flip() + 
  theme_minimal() + 
  labs  (x= "Tree Counts" , y = "legal Status")
  
```
**Example 2** Only going to keep observations 

```{r}
permitted_data_sf <- sf_trees %>% 
  filter(legal_status == "Permitted Site" & caretaker == "MTA")
```

**Example 3**

```{r}
blackwood_acacia_df <- sf_trees %>% 
  filter(species ==  "Acacia melanoxylon :: Blackwood Acacia") %>% #str_detect did not work for me :()
  select(legal_status, date , lat = latitude, lon = longitude)

#make a little graph of the locations 
ggplot(data = blackwood_acacia_df, aes(x = lon, y =lat)) + 
  geom_point(color = "darkgreen")

```
**Example 4** Use new function called separate 
```{r}
 sf_trees_separate <- sf_trees %>% 
  separate(species, into = c("spp_scientific" , "spp_common"), sep = "::")

```

**Example 5** Use unite 
```{r}
ex_5 <- sf_trees %>% 
  unite("id_status" , tree_id, legal_status, sep = "_COOL_")

```


**Part 2** Using `tmap`

###Step 1 - convert lat and lon to spatial points using `st_as_sf()`
```{r}
blackwood_acacia_sf <- blackwood_acacia_df %>% 
  drop_na(lat , lon) %>% 
  st_as_sf(coords= c('lon' , 'lat'))

### We need to tell R the coordinate reference system 
st_crs(blackwood_acacia_sf) <- 4326

ggplot(data = blackwood_acacia_sf) + 
  geom_sf(color = "darkgreen") + 
  theme_minimal()
```
### now that it is spatical we can add a map 
So read in SF shape file and add to map
```{r}
sf_map <- read_sf(here("data" , "sf_map" , "tl_2017_06075_roads.shp"))

sf_map_transform <- st_transform(sf_map, 4326)

ggplot(data = sf_map_transform) + 
  geom_sf()
```
```{r}
## Combining the maps 
ggplot() + 
  geom_sf(data = sf_map , 
          size = .1 ,
          color = "darkgrey") + 
  geom_sf(data = blackwood_acacia_sf, 
         color = "red", 
         size = 0.5) + 
  theme_void() 
```
### Now an interactive map! 

```{r}
tmap_mode("view")

tm_shape(blackwood_acacia_sf) + 
  tm_dots() # not working :()
```


